Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-01-24 paper/main.tex
Figures

In many practical optimization problems, the objective function is not
explicitly available and solutions are evaluated by means of expensive
prediction models, simulations or physical experiments. When decision variables
are continuous, the use of Bayesian surrogate-models (e.g., Gaussian processes)
is widespread in
optimization~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
success, there have been attempts at adapting such surrogate-based optimization
algorithms to the combinatorial case~\cite{MorKat2011:evo}, a notable example being Combinatorial
Efficient Global Optimization
(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
ruggedness of combinatorial landscapes, which makes local search particularly
effective in the white-box context, lessens the effectiveness of global
surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
expensive to train and optimize. The additional time required by the Bayesian
optimizer may impose a significant overhead in computation time, specially if
computation time of each function evaluation is measured in ``few'' minutes or
hours rather than days and when ``expensive'' refers to resources or economical
cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
ant colony optimization (ACO) %~\citep{DorStu2004:book}
is competitive with CEGO
on a black-box version of the travelling salesman problem under a budget of
$1\,000$ function evaluations.  On the other hand, $1\,000$ evaluations is
still a relatively large budget for CEGO, whereas ACO was not designed for such
short budgets.
%Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}.

@@

Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$. Otherwise, uBorda ranking will be closer to the rankings in $S$ with larger value of $w(\sigma)$. Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with high values of $w(\sigma$).
The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
%% MANUEL: To save space.
%For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
%
% In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$. Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
Finally, uBorda is computed in polynomial time.

@@

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are generated uniformly at random  (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to learn a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). 
Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$. 
A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. 
Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $\FEmax$.

@@

>>>>

In many practical optimization problems, the objective function is not
explicitly available and solutions are evaluated by means of expensive
prediction models, simulations or physical experiments. When decision variables
are continuous, the use of Bayesian surrogate-models (e.g., Gaussian processes)
is widespread in
optimization~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
success, there have been attempts at adapting such surrogate-based optimization
algorithms to the combinatorial case~\cite{MorKat2011:evo}, a notable example being Combinatorial
Efficient Global Optimization
(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
ruggedness of combinatorial landscapes, which makes local search particularly
effective in the white-box context, lessens the effectiveness of global
surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
expensive to train and optimize. The additional time required by the Bayesian
optimizer may impose a significant overhead in computation time, specially if
computation time of each function evaluation is measured in ``few'' minutes or
hours rather than days and when ``expensive'' refers to resources or economical
cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
ant colony optimization (ACO) %~\citep{DorStu2004:book}
is competitive with CEGO
on a black-box version of the travelling salesman problem under a budget of
$1\,000$ function evaluations.  On the other hand, $1\,000$ evaluations is
still a relatively large budget for CEGO, whereas ACO was not designed for such
short budgets.
%Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}. 
\manuel{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
 they mixed expensive to evaluate functions with black-box functions and of course they are different. In fact much of the metaheuristics do not mind
 if the problem is withe-box or black-box as they do not use static information from the function in the optimization process. In this case I understand
 that they have a black-box problem where the objective function is costly to evaluate. This should be clear and I would recommend to delete other comments
 related with white-box and black-box.
}{}

@@

Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$. Otherwise, uBorda ranking will be closer to the rankings in $S$ with larger value of $w(\sigma)$. Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with high values of $w(\sigma$).
The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
\ekhine{reesribe y ojo an grater / less}{}
%% MANUEL: To save space.
%For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
%
% In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$. Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
Finally, uBorda is computed in polynomial time.

@@

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are generated uniformly at random  (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to learn a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). 
Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$. 
A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. 
Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $\FEmax$.
\ekhine{(3) In section 3, second half of the second paragraph (from "Then, we sample a new candidate [...]"): the algorithm 1 is not properly described. Please provide a description in line with algorithm 1 and, above all, describe the rationale behind the choice of line 7 of algorithm 1.
I like it, but it deserves to be better described. It seems to me a sort of "novelty-driven" optimization.(4) A small detail, but requires a clarification. In page 6, learning step paragraph: the weight formula uses fitness in the exponent, while the base rho is in (0,1). Clearly the authors are working with minimization problems and want to increase the weight when fitness improves/decreases. This is ok, but this only happens with their formula when fitnesses are greater than 1. I know this is quite usual in the problems analyzed by the authors, but it may not be so usual in general. Please add a clarification.}{}

@@

\ekhine{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$. In my opinion the value of $\beta$ is quite relevant.}{}
